---
title: "MMPH6117 Final Examination"
author:
- 'Name: Mingze Gao'
- 'UID: 3035426241'
subtitle: Question 1
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

Read data sets.  
```{r Read Data}
library(readr)
exam1 <- read_csv("../Final/exam_question1.csv")
exam2 <- read_csv("../Final/exam_question2.csv")
exam3 <- read_csv("../Final/exam_question3.csv")
```

A) A  
```{r}
exam1$logcfr <- log(exam1$cfr)
exam1$se.logcfr <- (log(exam1$cfr.ub) - log(exam1$cfr.lb + 1e-6)) / (2 * 1.96)

library(metafor)
exam1.fe <- rma(yi = logcfr, sei = se.logcfr, slab = study, method = "FE", data = exam1)
exam1.fe

with(exam1.fe, exp(c(b, ci.lb, ci.ub)))
```
The overall estimate is 76.70 (95% CI = 75.08 - 78.35)  

B) B  
```{r}
Q <- exam1.fe$QE
I2 <- (Q - (exam1.fe$k - 1)) / Q * 100

I2
```
There is evidence of significant heterogeneity since the Higgins' I^2 is substantial high.  

C) C  
```{r}
exam1.re <- rma(yi = logcfr, sei = se.logcfr, slab = study, method = "REML", data = exam1)
exam1.re

with(exam1.re, exp(c(b, ci.lb, ci.ub)))
```
The overall estimate is 53.08 (95% CI = 44.27 - 63.65)  

D) D  
```{r}
forest(exam1.re, atransf = exp, refline = 0, at = log(c(0.1,10,100,500,1000)), showweights = T)
```

E) E  
```{r}
funnel(exam1.re, atransf=exp)
regtest(exam1.re)
```
The funnel plot shows no significant asymmetry and Egger's test null hypothesis is rejected since p > 0.01. So no strong evidence of publication bias.  

F) F  
```{r}
exam1$mark <- 0
exam1$mark[c(1,2,3,5,7,8,12,13,16,17,18,19)] <- 1

exam1.center <- exam1[exam1$mark==1,]
exam1.re.center <- rma(yi=logcfr, sei=se.logcfr, slab=study, method="REML", data=exam1.center)
with(exam1.re.center, exp(c(b, ci.lb, ci.ub)))

exam1.non <- exam1[exam1$mark==0,]
exam1.re.non <- rma(yi=logcfr, sei=se.logcfr, slab=study, method="REML", data=exam1.non)
with(exam1.re.non, exp(c(b, ci.lb, ci.ub)))
```
Based on the model applied to seperate regions, the hypothesis is correct as CFR from epicenters can be higher than non-epicenters with overall estimate 54.31 (95% CI: 41.80 - 70.56) compared with 50.37 (95% CI: 40.33 - 62.91)  

G) G  
1) The CFR of Covid-19 patients requiring IMV globally is quite high, after adjusting the heterogeneity of differenct studies, the overall CFR is 53.08.  
2) The CFR differs quite obvious between epicenters and non-epicenters, suggesting the overwhelmed healthcare system would deteriorate the situation.  
3) Although no stong publication bias was reported by funnel plot and following assessments, but some datasets contains small smple size which poses question on the validity of the resutllt.
4) The Olivares data has 0 on lower bound of CI which cannot perform log function, thus might not be credible and a 1e-6 is added to avoid that. Similarly, forest plot also assigned minimal weight on this data.  
