---
title: "MMPH6117 Final Examination"
author:
- 'Name: Mingze Gao'
- 'UID: 3035426241'
subtitle: Question 3
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

Read data sets.  
```{r Read Data}
library(readr)
exam1 <- read_csv("../Final/exam_question1.csv")
exam2 <- read_csv("../Final/exam_question2.csv")
exam3 <- read_csv("../Final/exam_question3.csv")
```

A) A  
```{r}
exam3$agegp <- 1
for(i in 1:nrow(exam3)) 
{
  if (exam3[i, "age"] >= 60)
  {
    exam3[i, "agegp"] = 2
  }
  else
  {
    if (exam3[i, "age"] <= 17)
    {
      exam3[i, "agegp"] = 0
    }
  }
}
round(with(exam3, prop.table(table(agegp, death),1)),3)
```
Agegp 0, 1, 2 corresponding to 0-17y, 18-59y and >= 60y respectively.  The the case fatality ratio by age groups are 0.6%, 1% and 38.1% respectively. A high fatality rate can be observed in elderly.  

B) B  
```{r}
require(MASS)
reb <- glm(death~age+comorbid, data=exam3, family=binomial)
summary(reb)
```
Based on the coefficients, increasing both age and comorbid could result in higher death rate, however, the p-value of comorbid for this model is not significant.  

C) C  
```{r}
exam3$age2 <- exam3$age * exam3$age
rec <- glm(death~age2+age+comorbid, data=exam3, family=binomial)
summary(rec)
```
Based on the coefficients, increasing both age^2 and comorbid could result in higher death rate whilst age poses negative effect on the outcome. However, the p-value of comorbid and age for this model is not significant. Also, the coefficients for age and age2 are very small, might not be credible.  

D) D 
```{r}
deviance(reb)/df.residual(reb)
deviance(rec)/df.residual(rec)
```
For reference, to evaluate satisfactory fitting of the model, the residual deviance / df should not be too much bigger than 1, better if it is below 1. The above model fits satisfactorily.  

E) E  
```{r}
AIC(reb, rec)
```
Regarding AIC metric, a lower AIC indicates a better model. So based on the results of AIC and goodness of fit, model C is preferred.  
However, model C is quadratic effect on age which is not necessary I suppose and since the difference between AIC and residual is small, model B might be better.  

F) F  
```{r}
require(car)
require(MASS)
library(survival)
vif(rec)
exam3$agescale <- scale(exam3$age, scale = F)
exam3$age2 <- exam3$agescale * exam3$agescale
ref <- glm(death~age2+age+comorbid, data=exam3, family=binomial)
summary(ref)
vif(ref)
```
The multicollinearity among the variables of model C is high for age and age2 because vif result is far away from 1.  
We could see that after certering the data, the vif can be reduced as well as the collinearity.  

G) G  
1) The crude fatality rate for zoonotic disease is higher for elderly people, whilst the difference is not tremendous between other adults and teenagers.  
2) Beside age, comorbid could also contribute to the death rate, however the effect might not be significant.  
3) Certering is sometims useful for tackling the collinearity problem. After reducing, both age2 and age are significant regarding p value. The effect of age2 is smaller than age.  